Limited physiological plasticity and reduced relative variance in physiological rates of ectotherm populations under climate change

### 2024-07-02 ms.qmd
Updated life history figure with stats. Need to check stats carefully as it required manual caluclation given fitted wasn't working. Did some checks and looks ok, but more warranted. Need to now update stats in results section for LH comparisons pooling, then clean up code

#######################################################################
# Bayesian posteriors from model 1.4
#######################################################################
 # Create the data we want to make predcitions for. Just more intuitive
  newdata <- data.frame( c_accl_time = 0, 
                          max_temp_c = 0, 
                                type = as.factor(rep(c("acute", "acclim"))), 
                             habitat = as.factor(rep(c("m", "f", "t"), each = 4)), 
                           life.hist = as.factor(c("j", "a", "a", "j")))
  newdata$type  <- relevel(newdata$type, ref = "acute")

@@

# Now we can do all the calculations in one shot.
         pred <- apply(t(posteriors_1_4), 2, function(x) as.matrix(X)%*%x)
     sum_pred <- apply(pred, 1, function(x) mean(x))
  sum_predL95 <- apply(pred, 1, function(x) quantile(x, 0.025))
  sum_predU95 <- apply(pred, 1, function(x) quantile(x, 0.975))

# Create a summary dataset. 
  newdata <- cbind(newdata, est = sum_pred, L = sum_predL95, u = sum_predU95)

# Now, lets calculate the contrasts between acute and acclim for a and j within each habitat. We need to use pred for that because it's the entire posterior distribution
    # Marine 
        m_adult <- pred[2,] - pred[3,] # Posterior contrast between acclimation and acute effects for adult, marine species
        m_jeuv <-  pred[4,] - pred[1,] # Posterior contrast between acclimation and acute effects for j, marine species
    # Freshwater
        f_ad <- 4
        f_adult <- pred[2+f_ad,] - pred[3+f_ad,] # Posterior contrast between acclimation and acute effects for adult, freshwater species
        f_jeuv <-  pred[4+f_ad,] - pred[1+f_ad,] # Posterior contrast between acclimation and acute effects for j, freshwater species
    # Terrestrial
        t_ad <- 8
        t_adult <- pred[2+t_ad,] - pred[3+t_ad,] # Posterior contrast between acclimation and acute effects for adult, terrestrial species
        t_jeuv <-  pred[4+t_ad,] - pred[1+t_ad,] # Posterior contrast between acclimation and acute effects for j, terrestrial species

# Now, lets calculate the contrasts between acute a and j as well as acclim a and j within each habitat. We need to use pred for that because it's the entire posterior distribution
    # Marine 
         m_acute_lh <-  pred[1,] - pred[3,] # Contrasts acute a and acute j values for marine species
        m_acclim_lh <-  pred[4,] - pred[2,] # Contrasts acclim a and acute j values for marine species
    # Freshwater
            f_ad <- 4
         f_acute_lh <- pred[1+f_ad,] - pred[3+f_ad,] # Contrasts acute a and acute j values for freshwater species
        f_acclim_lh <-  pred[4+f_ad,] - pred[2+f_ad,] # Contrasts acclim a and acute j values for freshwater species
    # Terrestrial
            t_ad <- 8
         t_acute_lh <- pred[1+t_ad,] - pred[3+t_ad,] # Contrasts acute a and acute j values for terrestrial species
        t_acclim_lh <-  pred[4+t_ad,] - pred[2+t_ad,]  # Contrasts acclim a and acute j values for terrestrial species

@@

@@

###########################################################
  # Processing Bayesian Posteriors
###########################################################
      # Create new data for predictions
       lh_b_data <- data.frame(     
                              max_temp_c = 0,
                             c_accl_time = 0,
                                    type = c("acclim", "acute"),
                                 habitat = rep(c("t", "m", "f"), each = 4),
                               life.hist = rep(c("j", "a"), each = 2))

       # Predict from model. Note this is the summary.
         pred_lncvr_lh_cvr <- fitted(model2.4_bayes, 
                                     newdata = lh_b_data, 
                                     re_formula = NA,
                                     summary = TRUE, 
                                     robust = FALSE)
         pred_lncvr_lh_cvr <- cbind(lh_b_data, pred_lncvr_lh_cvr)

       # Predict to get the posterior. Note that we have the full posterior distribution here for each of the 12 categories in the new data. Here, we want to look at the difference between acute and acclimation within each of the categories. 
            pred_lncvr_lh_post <- fitted(model2.4_bayes, 
                                     newdata = lh_b_data, re_formula = NA,
                                     summary = FALSE, 
                                     robust = FALSE)
        contrast_lncvr_lh_post <-  pred_lncvr_lh_post[,c(1,3,5,7,9,11)] - pred_lncvr_lh_post[,c(2,4,6,8,10,12)] # Acclimation - acute.

        # Contrast statistics
            contrast_lncvr_lh_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     trait = rep(c("", ""), each = 1),
                                                    diff = sapply(data.frame(contrast_lncvr_lh_post), mean),
                                                     l95 = sapply(data.frame(contrast_lncvr_lh_post), function(x) quantile(x, 0.025)),
                                                     u95 = sapply(data.frame(contrast_lncvr_lh_post), function(x) quantile(x, 0.975)),
                                                       p = sapply(data.frame(contrast_lncvr_lh_post), function(x) p_value(pmcmc(x))))

@@

## lnRR Q10

@@

p1 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "m"), group = "spp", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + labs(title = "Marine") + add_theme() + ylim(-0.4, 1.5) + theme(legend.position = "none", axis.text.y = element_blank()) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", m_r$k, " (", m_r$spp, ")")), size = size_annot)

p2 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Freshwater")  + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-0.4, 1.5) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", f_r$k, " (", f_r$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1)) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical"))

p3 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "t", legend.pos = "none"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Terrestrial") + add_theme()  +  ylim(-0.3, 1.8) + theme(legend.position = "none") + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", t_r$k, " (", t_r$spp, ")")), size = size_annot) + scale_x_discrete(labels = c("A" = "Adult","J" = "Juvenile"))

# CVR

m_cv <- n_lncvr_lh %>% filter(habitat == "m")
f_cv <- n_lncvr_lh %>% filter(habitat == "f")
t_cv <- n_lncvr_lh %>% filter(habitat == "t")

p1_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + ylim(c(-2, 2.2)) + add_theme() + ylim(-1.5, 1.1)  + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", m_cv$k, " (", m_cv$spp, ")")), size = size_annot) + theme(legend.position = "none", axis.text.y = element_blank())

p2_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type", angle = 45, g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)  + add_theme() + ylim(-1.3, 1.1) + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", f_cv$k, " (", f_cv$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1), axis.text.y = element_blank()) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical"))

p3_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "t"), angle = 45, group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)+ add_theme()  + ylim(-1.55, 1.1) + annotate("text", x = x.pos, y = 0.6, label = TeX(paste0("\\it{k} = ", t_cv$k, " (", t_cv$spp, ")")), size = size_annot) +  theme(legend.position = "none") + scale_x_discrete(labels = c("A" = "Adult","J" = "Juvenile"))

fig4 <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig4_lh_r1.pdf", width = 22, height = 12, device = "pdf")

>>>>

#######################################################################
# Bayesian posteriors from model 1.4
#######################################################################
 # Create the data we want to make predictions for. Just more intuitive
  newdata <- data.frame( c_accl_time = 0, 
                          max_temp_c = 0, 
                                type = relevel(factor(c("acclim", "acute")), ref = "acute"), 
                             habitat = rep(c("m", "f", "t"), each = 4), 
                           life.hist = rep(c("j", "a"), each = 2))

@@

# Now we can do all the calculations in one shot.
         pred <- apply(t(posteriors_1_4), 2, function(x) as.matrix(X)%*%x)
     sum_pred <- apply(pred, 1, function(x) mean(x))
  sum_predL95 <- apply(pred, 1, function(x) quantile(x, 0.025))
  sum_predU95 <- apply(pred, 1, function(x) quantile(x, 0.975))
            p <- apply(pred, 1, function(x) p_value(pmcmc(x)))

  # Create a summary dataset. 
  newdata <- cbind(newdata, est = sum_pred, L = sum_predL95, u = sum_predU95, p = p)

# Now, lets calculate the contrasts between acute and acclim for a and j within each habitat. We need to use pred for that because it's the entire posterior distribution

    # Contrast between Acclimation and Acute WITHIN each life-history stage. Double checked a bunch of calculations to make sure the right contrasts were being done. 
  contrast_lnrr_lh_post <-  pred[c(1,3,5,7,9,11),] - pred[c(2,4,6,8,10,12),] # Acclimation - acute.

        # Contrast statistics
            contrast_type_lnrr_lh_stats <- data.frame( habitat = rep(c("m", "f", "t"), each = 2),
                                                     stage = rep(c("j", "a"), each = 1),
                                                    diff = apply(data.frame(contrast_lnrr_lh_post), 1, mean),
                                                     l95 = apply(data.frame(contrast_lnrr_lh_post), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lnrr_lh_post), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lnrr_lh_post), 1, function(x) p_value(pmcmc(x))))

    # Contrast between Acclimation and Acute ACROSS each life-history stage. Double checked a bunch of comparisons.
          contrast_lnrr_lh_post_type <-  pred[c(4,3,8,7,12,11),] - pred[c(2,1,6,5,10,9),] # Adult - Jeu.

        # Contrast statistics
            contrast_type_lnrr_lh_stats <- data.frame( habitat = rep(c("m", "f", "t"), each = 2),
                                                     tyoe = rep(c("acute", "acclim"), each = 1),
                                                    diff = apply(data.frame(contrast_lnrr_lh_post_type), 1, mean),
                                                     l95 = apply(data.frame(contrast_lnrr_lh_post_type), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lnrr_lh_post_type), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lnrr_lh_post_type), 1, function(x) p_value(pmcmc(x))))
  ######## OLD CODE BELOW>
        m_adult <- pred[2,] - pred[3,] # Posterior contrast between acclimation and acute effects for adult, marine species
        m_jeuv <-  pred[4,] - pred[1,] # Posterior contrast between acclimation and acute effects for j, marine species

    # Freshwater
        f_ad <- 4
        f_adult <- pred[2+f_ad,] - pred[3+f_ad,] # Posterior contrast between acclimation and acute effects for adult, freshwater species
        f_jeuv <-  pred[4+f_ad,] - pred[1+f_ad,] # Posterior contrast between acclimation and acute effects for j, freshwater species

    # Terrestrial
        t_ad <- 8
        t_adult <- pred[2+t_ad,] - pred[3+t_ad,] # Posterior contrast between acclimation and acute effects for adult, terrestrial species
        t_jeuv <-  pred[4+t_ad,] - pred[1+t_ad,] # Posterior contrast between acclimation and acute effects for j, terrestrial species

# Now, lets calculate the contrasts between acute a and j as well as acclim a and j within each habitat. We need to use pred for that because it's the entire posterior distribution
    # Marine 
         m_acute_lh <-  pred[1,] - pred[3,] # Contrasts acute a and acute j values for marine species
        m_acclim_lh <-  pred[4,] - pred[2,] # Contrasts acclim a and acute j values for marine species

    # Freshwater
            f_ad <- 4
         f_acute_lh <- pred[1+f_ad,] - pred[3+f_ad,] # Contrasts acute a and acute j values for freshwater species
        f_acclim_lh <-  pred[4+f_ad,] - pred[2+f_ad,] # Contrasts acclim a and acute j values for freshwater species

    # Terrestrial
            t_ad <- 8
         t_acute_lh <- pred[1+t_ad,] - pred[3+t_ad,] # Contrasts acute a and acute j values for terrestrial species
        t_acclim_lh <-  pred[4+t_ad,] - pred[2+t_ad,]  # Contrasts acclim a and acute j values for terrestrial species

@@

   # Calculate the design matrix using this new data. Need to be careful here to check that the levels are set up the same as the posterior.

@@

###########################################################
  # Processing Bayesian Posteriors
###########################################################
      # Create new data for predictions
       lh_b_data <- data.frame( c_accl_time = 0, 
                          max_temp_c = 0, 
                                type = relevel(factor(c("acclim", "acute")), ref = "acute"), 
                             habitat = rep(c("m", "f", "t"), each = 4), 
                           life.hist = rep(c("j", "a"), each = 2))

            # I have no idea why the fitted function is not working on this model. We're doing this the old fashioned way
            X <- model.matrix(~max_temp_c + c_accl_time + type*habitat*life.hist, data = lh_b_data)

            # Extract posteriors
              post_lh <- posterior_samples(model2.4_bayes, pars = "^b")

        # Now we can do all the calculations in one shot.# Predict to get the posterior. Note that we have the full posterior distribution here for each of the 12 categories in the new data. Here, we want to look at the difference between acute and acclimation within each of the categories. 
            pred_lh <- apply(t(post_lh),2 , function(x) as.matrix(X)%*%x)

        contrast_lncvr_lh_post <-  pred_lh[c(1,3,5,7,9,11),] - pred_lh[c(2,4,6,8,10,12),] # Acclimation - acute.

        # Contrast statistics
            contrast_lncvr_lh_stats <- data.frame( habitat = rep(c("t", "m", "f"), each = 2),
                                                     stage = rep(c("j", "a"), each = 1),
                                                    diff = apply(data.frame(contrast_lncvr_lh_post), 1, mean),
                                                     l95 = apply(data.frame(contrast_lncvr_lh_post), 1, function(x) quantile(x, 0.025)),
                                                     u95 = apply(data.frame(contrast_lncvr_lh_post), 1, function(x) quantile(x, 0.975)),
                                                       p = apply(data.frame(contrast_lncvr_lh_post), 1, function(x) p_value(pmcmc(x))))
```

@@

###############
## lnRR Q10
###############

t_trait_b <- contrast_type_lnrr_lh_stats %>% filter(habitat == "t") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
m_trait_b <- contrast_type_lnrr_lh_stats %>% filter(habitat == "m") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
f_trait_b <- contrast_type_lnrr_lh_stats %>% filter(habitat == "f") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)

@@

p1 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "m"), group = "spp", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + labs(title = "Marine") + add_theme() + ylim(-0.4, 1.5) + theme(legend.position = "none", axis.text.y = element_blank()) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", m_r$k, " (", m_r$spp, ")")), size = size_annot) + annotate("text", x = yloc, y = -0.4, label = TeX(paste0("$\\beta$ = ", m_trait_b$diff, " [", m_trait_b$l95, " to ", m_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", m_trait_b$p)), hjust = 0, size = 8)

p2 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Freshwater")  + add_theme() + theme(legend.position = "none", axis.text.y = element_blank()) + ylim(-0.4, 1.5) + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", f_r$k, " (", f_r$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1)) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical")) + annotate("text", x = yloc, y = -0.4, label = TeX(paste0("$\\beta$ = ", f_trait_b$diff, " [", f_trait_b$l95, " to ", f_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", f_trait_b$p)), hjust = 0, size = 8)

p3 <- orchard_plot(model1.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "t", legend.pos = "none"), group = "record_num", xlab = TeX("$lnRR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size, angle = 45)+ labs(title = "Terrestrial") + add_theme()  +  ylim(-0.3, 1.8) + theme(legend.position = "none") + annotate("text", x = x.pos, y = 1.3, label = TeX(paste0("\\it{k} = ", t_r$k, " (", t_r$spp, ")")), size = size_annot) + annotate("text", x = yloc, y = -0.3, label = TeX(paste0("$\\beta$ = ", t_trait_b$diff, " [", t_trait_b$l95, " to ", t_trait_b$u95, "] ", "$\\,p_{MCMC}$ = ", t_trait_b$p)), hjust = 0, size = 8) +scale_x_discrete(labels = c("A" = "Adult","J" = "Juvenile"))

###############
# ln CVR
###############
m_cv <- n_lncvr_lh %>% filter(habitat == "m")
f_cv <- n_lncvr_lh %>% filter(habitat == "f")
t_cv <- n_lncvr_lh %>% filter(habitat == "t")

  t_trait_cvr <- contrast_lncvr_lh_stats %>% filter(habitat == "t") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
m_trait_b_cvr <- contrast_lncvr_lh_stats %>% filter(habitat == "m") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)
f_trait_b_cvr <- contrast_lncvr_lh_stats %>% filter(habitat == "f") %>% purrr::modify_if(~is.numeric(.), ~round(., 2)) %>% arrange(stage)

p1_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "m"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, angle = 45, branch.size = branch.size, trunk.size = trunk.size) + ylim(c(-2, 2.2)) + add_theme() + ylim(-1.5, 1.1)  + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", m_cv$k, " (", m_cv$spp, ")")), size = size_annot) + theme(legend.position = "none", axis.text.y = element_blank()) + annotate("text", x = yloc, y = -1.3, label = TeX(paste0("$\\beta$ = ", m_trait_b_cvr$diff, " [", m_trait_b_cvr$l95, " to ", m_trait_b_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", m_trait_b_cvr$p)), hjust = 0, size = 8)

p2_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "f"), group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type", angle = 45, g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)  + add_theme() + ylim(-1.3, 1.1) + annotate("text", x = x.pos, y = 0.8, label = TeX(paste0("\\it{k} = ", f_cv$k, " (", f_cv$spp, ")")), size = size_annot) + theme(legend.position = "right", legend.justification = c(1, 1), axis.text.y = element_blank()) + guides(shape = guide_legend(direction = "vertical"), size = guide_legend(direction = "vertical")) + annotate("text", x = yloc, y = -1.3, label = TeX(paste0("$\\beta$ = ", f_trait_b_cvr$diff, " [", f_trait_b_cvr$l95, " to ", f_trait_b_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", f_trait_b_cvr$p)), hjust = 0, size = 8)

p3_cvr <- orchard_plot(model2.4, mod = "life.hist", at = list(c_accl_time = 0, habitat = "t"), angle = 45, group = "record_num", xlab = TeX("$lnCVR_{Q_{10}}$"), condition.lab = "Type", by = "type",  g = FALSE, k = FALSE, branch.size = branch.size, trunk.size = trunk.size)+ add_theme()  + ylim(-1.55, 1.1) + annotate("text", x = x.pos, y = 0.6, label = TeX(paste0("\\it{k} = ", t_cv$k, " (", t_cv$spp, ")")), size = size_annot) +  theme(legend.position = "none") + annotate("text", x = yloc, y = -1.55, label = TeX(paste0("$\\beta$ = ", t_trait_cvr$diff, " [", t_trait_cvr$l95, " to ", t_trait_cvr$u95, "] ", "$\\,p_{MCMC}$ = ", t_trait_cvr$p)), hjust = 0, size = 8) + scale_x_discrete(labels = c("A" = "Adult","J" = "Juvenile"))

fig4 <-(p3 | p1 | p2) / (p3_cvr | p1_cvr | p2_cvr) + plot_annotation(tag_levels = "A") 
ggsave(fig4, filename = "./output/figures/fig4_lh_r1.pdf", width = 28, height = 14, device = "pdf")
